Presentation Title

### 2017-06-21 thesis/chapters/ch2-model.tex
Write more related work
Change model to pure async

The primary goal guiding our design is scalability.
As mentioned in the Introduction, having a scalable blockchain system while still keeping global consensus
allows the system to be ubiquitous and realise the full potential of blockchain.

The secondary goal is to design an application neutral system.
In particular, it should act as a framework that provides the building blocks of blockchain based applications.
Application developers using the framework should be able to create any application they wish.
Further, we do not impose on a consensus algorithm, as long as it satisfies the properties of atomic broadcast which we describe in~\Cref{sec:overview-cons}.

Due to the nature of our system, we do not explicitly address the Sybil attack~\cite{douceur2002sybil}.
Sybil defence mechanism always require some form of reputation score from the application.
For example, social network based Sybil defence mechanims use graph structure of real-world relationships~\cite{yu2006sybilguard}.
Online marketplaces such as Amazon use the rating of buyer and sellers.
Thus it is not possible to design a Sybil defence mechanism with a a application neutral framework.
On the other hand, our system also has no restrictions on the Sybil defence technique
and application designers can pick the best mechanism for their application.

The third and final goal is security.
Our system should be unaffected in the presence of powerful adversaries.
Security is often difficult to verify, especially when it is not formalised, therefore we require our design to be provably secure.
To summarise, our system design is designed with the following goals in mind.
\begin{itemize}
    \item Application neutrality,
    \item scalability and
    \item security.
\end{itemize}

@@

\subsection{Extended TrustChain}
Extended TrustChain naturally builds on top of TrustChain, thus we first describe the standard TrustChain.
Our description has minor differences compared to the description in~\cite{trustchain}.
This is to help with the description of the extended TrustChain.
However, the two descriptions are functionally the same.

\subsubsection*{Standard TrustChain}
In TrustChain, every node has a ``personal'' chain. 
Initially, the chain only contains a genesis block.
When a node wishes to add a new transaction (TX), a new TX block is generated and is appended to the chain.
A TX block must have a valid hash pointer pointing to the previous block
and a reference\footnote{This is different from the original TrustChain definition found in~\cite{trustchain}.
In there, a TX block has two outgoing edges which are hash pointers to the two parties involved in the transaction.
This work uses one outgoing edge and a reference.} to its \emph{pair}.
As a result, a single transaction generates two TX blocks, one on each party's chain.
An example of is shown in \Cref{fig:trustchain-bad}.

@@

If every node follows the rules of TrustChain and we only consider hash pointers,
then the chain effectively forms a singly linked list.
However, if a node violates the rules, then a \emph{fork} may happen.
That is, there may be more than one TX block with a hash pointer pointing back to the same block.
In \Cref{fig:trustchain-bad}, node $b$ (in the middle chain) created two TX blocks that both point to $t_{b, 5}$.
If this is a ledger system it can be seen as a double spend, where the currency accumulated up until $t_{b, 5}$ are spent twice.

\subsubsection*{Extended TrustChain}
We are now ready to explain the Extended TrustChain, which we abbreviate to ETC.
In ETC, we introduce a new type of block---checkpoint (CP) block.
In contract to TX blocks, CP blocks do not store transactions or contain references.
Their purpose is to capture the state of the chain and the state of the whole system.
In particular, the state of the chain is captured with a hash pointer.
The state of the whole system is captured in the content of the CP block,
namely as a digest of the latest \emph{consensus result} which we explain in \Cref{sec:overview-cons}.
A visual representation is shown in \Cref{fig:trustchain-bad-cp}.

@@

% The consensus protocol uses an existing Byzantine consensus algorithm (explained in detail in \Cref{sec:xyz}) and runs it continuously in rounds.
The consensus protocol can be seen as a technique of running infinitly many rounds of some 
Byzantine consensus algorithm\footnote{More accuratly it is ACS or asynchronous subset consensus, we describe ACS in \Cref{sec:acs}.},
starting a new execution immediately after the previous one is completed.
This is necessary because blockchain systems always need to reach consensus on new values, or CP blocks in our case.

The high message complexity of Byzantine consensus algorithms prohibits us from running it on a large network.
Thus, for every round, we randomly select some node---called facilitators---to collect CP blocks and use them as the proposal.
The facilitators are elected using a \emph{luck value}, which is computed using $H(\C_r || pk_i)$,
where $\C_r$ is the consensus result (a set of CP blocks) in round $r$ and $pk_i$ is the public key of $i$.
Intuitively, the election is guaranteed to be random 
because the output of a cryptographically secure hash function is unpredictable and $\C_r$ cannot be determined in advance.

A visual explaination can be found in \Cref{app:consensus-example},
it walks through the steps needed for a node to be selected as a facilitator.

@@

The consensus and transaction protocol by themselves do not provide a mechanism to detect forks or other forms of tamperaing.
Thus we need a validation protocol to counteract malicious behaviour.
When a node wish to validate one of its TX, it asks the counterparty for the \emph{fragment} of the TX.
A fragment of a TX is a section of the chain beginning and ending with CP blocks that contains the TX.
Upon the counterparty's response, the node checks that the CP blocks are in consensus, the hash pointers are valid and his TX is actually in the fragment.
The TX is valid if these conditions are satisfied.
Intuitively, this works because it is hard (because hash collision is hard)
 to create a different chain that begins and ends with the same two CP blocks but with a different middle section.

@@

Our subprotocol design gives us the highly desireable non-blocking property.
In particular, we do not need to ``freeze'' the state of the chain for some communication to complete in order to create a block.
For instance, a node may start the consensus protocol, and while it is running, the node may still perform transactions.
By the time the consensus protocol is done, the new CP block is added to whatever the state that the chain is in.
It is not necessary to keep the chain immutable while the consensus protocol is running.

@@

For notational clarify, we use the following convention (adapted from~\cite{miller2016honey}) throughout this work.
\begin{itemize}
\item Lower case (e.g. $x$) denotes a scalar object or a tuple.
\item Upper case (e.g. $X$) denotes a set or a constant.
\item Sans serif (e.g. $\textsf{fn}(\cdot)$) denotes a function.
\item Typewrite (e.g. $\texttt{ack}$) denotes message type.
\end{itemize}
Further, we use $a || b$ to dedote concatination of the binary representations of $a$ and $b$.

We assume a distributed network where nodes are fully connected,
the channels are reliable\footnote{
    Reliability can be achieved in unreliable networks by resending messages or using some error correction code.
},
but messages may be re-ordered and delayed by at most some time $\Delta$,
this is sometimes known as a $\Delta$-synchronous network.
We assume there exist a Public Key Infrastructure (PKI), and nodes are identified by their unique and permanent public key.
Finally, we use the random oracle model, i.e. calls to the random oracle are denoted by $\textsf{H}: \{0, 1\}^* \rightarrow \{0, 1\}^\lambda$,
where $\{0, 1\}^*$ denotes the space of finite binary strings and $\lambda$ is the security parameter~\cite{bellare1993random}.

In our model we consider $N$ nodes, which is the population size.
$n$ of them are facilitators, $t$ out of $n$ are malicious and the inequality
$n \ge 3t + 1$ must hold.

We use a restricted version of the adaptive corruption model.
The first restriction is that corrupted node can only change across rounds.
That is, if a round has started, the corrupted nodes cannot changed until the next round.
The second restriction is that the adversary, presumably controlling all the corrupted nodes, is forgetful.
Namely the adversary may learn the internal state such as the private key of a corrupted node,
but if the node recovers, then the adversary must forget the private key.
Otherwise the adversary can eventually learn all the private keys and sabotage the system.
Finally, we assume computational security.
That is, the adversary can run polynomial-time algorithms but not exponential-time algorithms.

The primary data structure used in our system is Extended TrustChain.
Each node $u$ has a public and private key pair---$pk_u$ and $sk_u$, and a chain $B_u$.
The chain consist of blocks $B_u = \{ b_{u, i} : k \in \{ 0, \dots, h - 1 \} \}$,
where $b_{u, i}$ is the $i$th block of $u$,
and $h = |B_u|$.
We often use $b_{u, h}$ to denote the latest block.
There are two types of blocks, TX blocks and CP blocks.
If $T_u$ is the set of TX blocks of $u$ and $C_u$ is the set of CP blocks of $u$,
then it must be the case that $T_u \cup C_u = B_u$ and $T_u \cap C_u = \varnothing$.
The notation $b_{u, i}$ is generic over the block type.
We assume there exist a function $\textsf{typeof}: B_u \rightarrow \{ \tau, \gamma \}$ that returns the type of the block,
where $\tau$ represents the TX type and $\gamma$ represents the CP type.

\subsection{TX Block}
The TX block is a six-tuple, i.e $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$.
We describe each item in turn.
\begin{enumerate}
\item $\textsf{H}(b_{u, i - 1})$ is the hash pointer to the previous block.
\item $seq_u$ is the sequence which should equal $i$.
\item $txid$ is a cryptographically secure random number representing the transaction identifier.
\item $pk_v$ is the public key of the counterparty.
\item $m$ is the transaction message.
\item $sig_u$ is the signature created using $sk_u$ on the concatination of the binary representation of the five items above.
\end{enumerate}
The fact that we have no constraint on the content of $m$ is in alignment with our design goal---application neutrality.

TX blocks come in pairs.
In particular, for every block $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$
there exist one and only one \emph{pair} $t_{v, j} = \langle \textsf{H}(b_{v, j - 1}), seq_v, txid, pk_u, m, sig_v \rangle$.
Note that the $txid$ and $m$ are the same, and the public keys refer to each other.
Thus, given a TX block, these properties allow us to identify its pair.

@@

\subsection{CP Block}

The CP block is a five-tuple, 
i.e. $c_{u, i} = \langle \textsf{H}(b_{u, i-1}), seq_u, \textsf{H}(\C_r), r, sig_u \rangle$,
where $\C_r$ is the consensus result in round $r$, the other items are the same as the TX block definition.
Note that unlike in our prior work~\cite{implicitconsensus}, CP blocks and TX blocks do not have independent sequence numbers.

The genesis block in the chain must be a CP block in the form of
$c_{u, 0} = \langle \textsf{H}(\bot), 0,  \textsf{H}(\bot), 0, sig_u \rangle$
where $\textsf{H}(\bot)$ can be interpreted as applying the hash function on an empty string.
The genesis block is unique due to every node due to $sig_u$.

\subsection{Consensus Result}
Our consensus protocol runs in rounds as discussed in \Cref{sec:system-overview}.
Every round is identified by a round number $r$, which is incremented on every new round.
The consensus result is a tuple, i.e. $\C_r = \langle r, C \rangle$,
where $C$ is a set of CP blocks agreed by the facilitators of round $r$.

@@

If there exist a tuple $\langle c_{u,a}, c_{u, b} \rangle$ for a TX block $t_{u, i}$,
where 
$$a = \argmin_{k, k < i, \texttt{typeof}(b_{u,k}) = \gamma}(i - k)$$
$$b = \argmin_{k, k > i, \texttt{typeof}(b_{u,k}) = \gamma}(k - i),$$
then $\langle c_{u,a}, c_{u, b} \rangle$ is the \emph{enclosure} of $t_{u, i}$.
Some TX blocks may not have any subsequent CP blocks, then its enclosure is $\bot$.

If the enclosure of some TX block is $\langle c_{u,a}, c_{u, b} \rangle$,
then its \emph{fragment} is computed as $\{ b_{u, i} : a \le i \le b \}$.
For convenience, the function $\textsf{fragment}(\cdot)$ represents the fragment of some TX block if it exists, otherwise $\bot$.

\emph{Agreed enclosure} is the same as enclosure with an extra constraint where the CP blocks must be in some consensus result $\C_r$ and also must be the smallest enclosure.
That is, suppose a chain is in the form
    $\{c_{i}, c_{i+1}, t_{i+2}, c_{i+3}\}$\footnote{Usually the notation is of the form $c_{u, i}$, but the node identity is not important here so we simplify it to $c_{i}$}
    and $c_{i}, c_{i+1}, c_{i+3}$ are in $\C_r, \C_{r+1}, \C_{r+3}$ respectively,
    then the agreed enclosure of $t_{i+2}$ is $\langle c_{i+1}, c_{i+3}\rangle$ and cannot be $\langle c_{i}, c_{i+3}\rangle$.
Similarly, \emph{agreed fragment} is computed using the agreed enclosure.
We define its function to be $\textsf{agreed\_fragment}(\cdot)$

The length of the fragment is constrained by $L$,
namely $\forall t |\textsf{fragment}(t)| \le L$.
The purpose to prevent spam and encourage nodes to create more CP blocks.
$L$ should be sufficiently high so that busy nodes are not hindered by it.

@@

Our consensus protocol runs on top of the model described above.
It is directly related to the creation of CP blocks.
The objectives of the protocol are to
    allow honest nodes always make progress (by creating new CP blocks),
    compute correct consensus result in every round
    and have unbiased election of facilitators.
Concretely, we define the necessary properties as follows,
it is collectively known as the correctness properties of the consensus protocol.
\begin{definition}
\label{def:consensus}
\textbf{\emph{Correctness properties of the consensus protocol}}

$\forall r \in \mathbb{N}$, the following properties must be satisfied.
\begin{enumerate}
    \item \emph{Agreement}:
        If one honest node outputs a list of facilitators $\F_r$,
        every other node honest decides on $\F_r$
    \item \emph{Validity}:
        If any honest node outputs $\F_r$, then $|\F_r| > n$ and $|\F_r|$ must contain at least $n - t$ honest nodes.
    \item \emph{Fairness}:
        Every node should have an equal probability of becoming a facilitator.
    \item \emph{Termination}:
        At the end of the round, every node outputs some $\F_r$.
\end{enumerate}

\end{definition}
Note that the final three properties are on the facilitators rather than on every node,
they are the properties ACS. 
The fair lottery and consistent facilitator are properties on every node,
they are prerequites for ACS.

Before describing the protocol in detail, we take a brief detour to give background on the Byzantine consensus algorithm we use in the consensus protocol.
% Starting with the bootstrap phase and then moving on to the actual consensus phase.

\subsection{Background on Asynchronous Subset Consensus}

The best way to explain asynchronous subset consensus (ACS) is to contrast it with the typical Byzantine consensus problem.
We adapt the description from~\cite{podc}.
%The Byzantine consensus problem is a classical problem is distributed systems literature.
%It is first described by Leslie Lamport as the Byzantine general's problem in 1982~\cite{lamport1982byzantine}.
%Byzantine consensus is described as follows (adapted from ).
\begin{definition}
\textbf{\emph{Byzantine Consensus}}

There are $n$ nodes, of which at most $t$ might experience Byzantine fault.
Node $i$ starts with an input value $v_i$.
The nodes must decide for one of  those values, satisfying the the following.
\begin{enumerate}
    \item \emph{Agreement:}
        All correct nodes decide for the same value.
    \item \emph{Validity:}
        The decision value must be the input value of a node.
    \item \emph{Termination:}
        All correct nodes terminate in finite time.
\end{enumerate}
\end{definition}
A node under Byzantine fault means that it can have arbitrary behaviour.
For example not sending message or colluding with other Byzantine nodes to undermine the entire system.
Note that the decision is on a single value.
This is in contrast to ACS which we describe next.

@@

There are $n$ nodes, of which at most $t$ might experience Byzantine fault.
Node $i$ starts with an non-empty set of input values $C_i$.
The nodes must decide an output $C$, satisfying the following.
\begin{enumerate}
    \item \emph{Validity:}
        If any correct party outputs a set $C$,
        then $|C| \ge n - t$ and $C$ contains the input of at least $n - 2t$ parties.
    \item \emph{Agreement:}
        If a correct party outputs $C$, then every party outputs $C$.
    \item \emph{Totality:}
        If $n - t$ party receive an input, then all correct parties produce an output.
\end{enumerate}
\end{definition}
ACS has the nice property of censorship resiliance when compared to other consensus algorithms.
For instance, Hyperledger and Tender mint uses Practical Byzantine Fault Tolerance (PBFT) as their consensus algorithm.
In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectivly filter transactions.
In contract, every party in ACS are involved in the proposal phase,
and it guarantees that if $n - 2t$ parties propose the same transaction, then it must be in the agreed output.
Thus, some value is submitted to at least $n - 2t$ nodes, it is guaranteed to be in the consensus result.
For a detailed description of ACS we refer to the HoneyBadgerBFT work~\cite{miller2016honey}.

The main drawback with ACS and also Byzantine consensus algorithms is the high message complexity.
Typically, such protocols have the complexity of $O(n^2)$, where $n$ is the number of parties.
Hence, it may work with a small number of nodes,
but it is infeasible for blockchain systems where thousands of nodes are involved.

@@

\subsection{Bootstrap Phase}
\label{sec:bootstrap}
Recall that facilitators are computed from the consensus result,
but the consensus result is agreed by the facilitators.
Thus we have a dependency cycle.
The goal of the bootstrap phase is to give us a starting point in the cycle.

Imagine that there is some bootstrap oracle, that initiates the code on every node.
The code satisfied all the properties in \Cref{def:consensus}.
Namely every node has the same set of valid facilitators $\F_1$ that are randomly choosen.
This concludes the bootstrap phase.

@@

@@

    \State $txid', pk'_u, m' \gets t_{v, j}$
    \If{$m \ne m' \vee pk_u \ne pk'_u \vee |F_{v, j}| > L$}
        \State \Return \emph{invalid}
    \EndIf \Comment no tampering
    \State

>>>>

The goals guiding our system design are as follows.
\begin{itemize}
    \item Performance and scalability,
    \item application neutrality and
    \item security.
\end{itemize}

As mentioned in the Introduction, the primary goal is performance and scalability.
Having a scalable blockchain system while still keeping global consensus
allows the system to be ubiquitous and realise the full potential of blockchain.

The secondary goal is to design an application neutral system.
In particular, it should act as a backbone that provides the building blocks of blockchain based applications.
It should be possible to buld any kind of application on top our system.
Further, we do not impose on a consensus algorithm, as long as it satisfies the properties of atomic broadcast which we describe in~\Cref{sec:overview-cons}.

Due to the nature of our system, we do not explicitly address the Sybil attack~\cite{douceur2002sybil}.
Sybil defence mechanism always require some form of reputation score from the application.
For example, social network based Sybil defence mechanims use graph structure of real-world relationships~\cite{yu2006sybilguard}.
Online marketplaces such as Amazon use the rating of buyer and sellers.
Thus it is not possible to design a Sybil defence mechanism with a application neutral framework.
On the other hand, our system also has no restrictions on the Sybil defence technique
and applications built on top of our system can use the best mechanism for their purpose.

The third and final goal is security.
Our system should be unaffected in the presence of powerful adversaries.
In particular, adversaries are Byzantine meaning that they can have arbitrary behaviour.
Thus anything is possible from simply ommitting messages to colluding with eachother in order to undermine the whole system.

@@

\subsection{Extended TrustChain}
Extended TrustChain naturally builds on top of the standard TrustChain. 
Thus we first describe the standard TrustChain.
Our description has minor differences compared to the description in~\cite{trustchain}.
This is to help with the description of the extended TrustChain.
We remark the difference when it occurs.
However, the two descriptions are functionally the same.

\subsubsection*{Standard TrustChain}
In TrustChain, every node has a ``personal'' chain. 
Initially, the chain only contains a genesis block generated by the nodes themselves.
When a node $A$ wishes to add a new transaction (TX) with $B$, a new TX block is generated and then appended to $A$'s chain.
The TX block must have a valid hash pointer pointing to the previous block
and a reference\footnote{This is different from the original TrustChain definition found in~\cite{trustchain}.
In there, a TX block has two outgoing edges which are hash pointers to the two parties involved in the transaction.
This work uses one outgoing edge and a reference.} to its \emph{pair} on $B$'s chain.
As a result, a single transaction generates two TX blocks, one on each party's chain.
An example of is shown in \Cref{fig:trustchain-bad}.

@@

If every node follows the rules of TrustChain and we only consider hash pointers,
then every chain effectively forms a singly linked list.
However, if a node violates the rules, then a \emph{fork} may happen.
That is, there may be more than one TX block with a hash pointer pointing back to the same block.
In \Cref{fig:trustchain-bad}, node $b$ (in the middle chain) created two TX blocks that both point to $t_{b, 5}$.
If this is a ledger system it can be seen as a double spend, where the currency accumulated up until $t_{b, 5}$ are spent twice.

\subsubsection*{Extended TrustChain}
We are now ready to explain the Extended TrustChain.
The primary difference is the introduction of a new type of block---checkpoint (CP) block.
In contract to TX blocks, CP blocks do not store transactions or contain references.
Their purpose is to capture the state of the chain and the state of the whole system.
In particular, the state of the chain is captured with a hash pointer.
The state of the whole system is captured in the content of the CP block,
namely as a digest of the latest \emph{consensus result} which we explain in \Cref{sec:overview-cons}.
A visual representation is shown in \Cref{fig:trustchain-bad-cp}.

From this point onwards, we use TrustChain to mean the Extended TrustChain unless explicitly clarified.

@@

% The consensus protocol uses an existing Byzantine consensus algorithm (explained in detail in \Cref{sec:xyz}) and runs it continuously in rounds.
The consensus protocol can be seen as a technique of running infinitly many rounds of some 
Byzantine consensus algorithm\footnote{More accuratly it is ACS or asynchronous subset consensus, we describe ACS in \Cref{sec:acs-background}.},
starting a new execution immediately after the previous one is completed.
This is necessary because blockchain systems always need to reach consensus on new values proposed by nodes in the system,
or CP blocks in our case.

The high communication complexity of Byzantine consensus algorithms prohibits us from running it on a large network.
Thus, for every round, we randomly select some node---called facilitators---to collect CP blocks from every other node and use them as the proposal.
The facilitators are elected using a \emph{luck value}, which is computed using $H(\C_r || pk_u)$,
where $\C_r$ is the consensus result 
(which can be seen as the set union of all the CP blocks collected by the facilitators)
in round $r$ and $pk_u$ is the public key of node $u$.
Intuitively, the election is guaranteed to be random 
because the output of a cryptographically secure hash function is unpredictable and $\C_r$ cannot be determined in advance.

A visual explaination can be found in \Cref{app:consensus-example},
it walks through the steps needed for a node to be selected as a facilitator using an example.

@@

The consensus and transaction protocol by themselves do not provide a mechanism to detect forks or other forms of tamperaing.
Thus we need a validation protocol to counteract malicious behaviour.
When a node wish to validate one of its TX, it asks the counterparty for the \emph{fragment} of the TX.
A fragment of a TX is a section of the chain beginning and ending with CP blocks that contains the TX and are in consensus.
Upon the counterparty's response, the node checks that the CP blocks are indeed in consensus,
the hash pointers are valid and his TX is actually in the fragment.
The TX is valid if these conditions are satisfied.
Intuitively, this works because it is hard (because a cryptographically secure hash function is second-preimage resistant)
to create a different chain that begins and ends with the same two CP blocks but with a different middle section.

@@

Our subprotocol design gives us the highly desireable non-blocking property.
In particular, we do not need to ``freeze'' the state of the chain for some communication to complete in order to create a block.
For instance, a node may start the consensus protocol, and while it is running, create new transactions and validate old transactions.
By the time the consensus protocol is done, the new CP block is added to whatever the state that the chain is in.
It is not necessary to keep the chain immutable while the consensus protocol is running.

@@

This section and the ones following it give a technical treatment of what the content in System Overview.
For notational clarity, we use the following convention (adapted from~\cite{miller2016honey}) throughout this work.
\begin{itemize}
\item Lower case (e.g. $x$) denotes a scalar object or a tuple.
\item Upper case (e.g. $X$) denotes a set or a constant.
\item Sans serif (e.g. $\textsf{fn}(\cdot)$) denotes a function.
\item Typewrite (e.g. $\texttt{ack}$) denotes message type.
\end{itemize}
Further, we use $a || b$ to denote concatination of the binary representations of $a$ and $b$.

We assume purely asynchronous channels with eventual delivery.
Thus in no circumstance do we make timing assumptions.
The adversary has fully control of the delivery schedule and the message ordering.
But they are not allowed to drop messages\footnote{
    Reliability can be achieved in unreliable networks by resending messages or using some error correction code.
}.

We assume there exist a Public Key Infrastructure (PKI), and nodes are identified by their unique and permanent public key.
Finally, we use the random oracle model, i.e. calls to the random oracle are denoted by $\textsf{H}: \{0, 1\}^* \rightarrow \{0, 1\}^\lambda$,
where $\{0, 1\}^*$ denotes the space of finite binary strings and $\lambda$ is the security parameter~\cite{bellare1993random}.

In our model we consider $N$ nodes, which is the population size.
$n$ of them are facilitators, $t$ out of $n$ are malicious and the inequalities
$n \ge 3t + 1$ and $N \ge n + t$ must hold.
The first is from the FLP impossibility result TODO.
The second is necessary due to our system design, which becomes clear in TODO.

Our threat model is as follows. 
We use a restricted version of the adaptive corruption model.
The first restriction is that corrupted node can only change across rounds.
That is, if a round has started, the corrupted nodes cannot changed until the next round.
The second restriction is that the adversary, presumably controlling all the corrupted nodes, is forgetful.
Namely the adversary may learn the internal state such as the private key of a corrupted node,
but if the node recovers, then the adversary must forget the private key.
Otherwise the adversary can eventually learn all the private keys and sabotage the system.
Finally, we assume computational security.
That is, the adversary can run polynomial-time algorithms but not exponential-time algorithms.

\section{Extended TrustChain}
The primary data structure used in our system is the Extended TrustChain.
Each node $u$ has a public and private key pair---$pk_u$ and $sk_u$, and a chain $B_u$.
The chain consist of blocks $B_u = \{ b_{u, i} : k \in \{ 0, \dots, h - 1 \} \}$,
where $b_{u, i}$ is the $i$th block of $u$,
and $h = |B_u|$.
We often use $b_{u, h}$ to denote the latest block.
There are two types of blocks, TX blocks and CP blocks.
If $T_u$ is the set of TX blocks of $u$ and $C_u$ is the set of CP blocks of $u$,
then it must be the case that $T_u \cup C_u = B_u$ and $T_u \cap C_u = \varnothing$.
The notation $b_{u, i}$ is generic over the block type.
We assume there exist a function $\textsf{typeof}: B_u \rightarrow \{ \tau, \gamma \}$ that returns the type of the block,
where $\tau$ represents the TX type and $\gamma$ represents the CP type.

\subsection{Transaction Block}
The TX block is a six-tuple, i.e $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$.
We describe each item in turn.
\begin{enumerate}
\item $\textsf{H}(b_{u, i - 1})$ is the hash pointer to the previous block.
\item $seq_u$ is the sequence number which should equal $i$.
\item $txid$ is the transaction identifier, it should be generated using a cryptographically secure random number generator by the initiator of the transaction.
\item $pk_v$ is the public key of the counterparty $v$.
\item $m$ is the transaction message.
\item $sig_u$ is the signature created using $sk_u$ on the concatination of the binary representation of the five items above.
\end{enumerate}
The fact that we have no constraint on the content of $m$ is in alignment with our design goal---application neutrality.

TX blocks come in pairs.
In particular, for every block 
$$t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$$
there exist one and only one \emph{pair} 
$$t_{v, j} = \langle \textsf{H}(b_{v, j - 1}), seq_v, txid, pk_u, m, sig_v \rangle.$$
Note that the $txid$ and $m$ are the same, and the public keys refer to each other.
Thus, given a TX block, these properties allow us to identify its pair.

@@

\subsection{Checkpoint Block}

The CP block is a five-tuple, 
i.e. $c_{u, i} = \langle \textsf{H}(b_{u, i-1}), seq_u, \textsf{H}(\C_r), r, sig_u \rangle$,
where $\C_r$ is the consensus result (which we describe in \Cref{sec:consensus-result}) in round $r$, the other items are the same as the TX block definition.
Note that unlike in our prior work~\cite{implicitconsensus}, CP blocks and TX blocks do not have independent sequence numbers.

The genesis block in the chain must be a CP block in the form of
$c_{u, 0} = \langle \textsf{H}(\bot), 0,  \textsf{H}(\bot), 0, sig_u \rangle$
where $\textsf{H}(\bot)$ can be interpreted as applying the hash function on an empty string.
The genesis block is unique because every node has a unique public and private key pair.

\subsection{Consensus Result}
\label{sec:consensus-result}
Our consensus protocol runs in rounds as discussed in \Cref{sec:system-overview}.
Every round is identified by a round number $r$, which is incremented on every new round.
The consensus result is a tuple, i.e. $\C_r = \langle r, C \rangle$,
where $C$ is a set of CP blocks agreed by the facilitators of round $r$.

@@

If there exist a tuple $\langle c_{u,a}, c_{u, b} \rangle$ for a TX block $t_{u, i}$,
where 
$$a = \argmin_{k, k < i, \texttt{typeof}(b_{u,k}) = \gamma}(i - k)$$
$$b = \argmin_{k, k > i, \texttt{typeof}(b_{u,k}) = \gamma}(k - i),$$
then $\langle c_{u,a}, c_{u, b} \rangle$ is the \emph{enclosure} of $t_{u, i}$.
Note that $c_{u, a}$ is the more recent CP block.
Also, some TX blocks may not have any subsequent CP blocks, then its enclosure is $\bot$.

If the enclosure of some TX block is $\langle c_{u,a}, c_{u, b} \rangle$,
then its \emph{fragment} is defined as $\{ b_{u, i} : a \le i \le b \}$.
For convenience, the function $\textsf{fragment}(\cdot)$ represents the fragment of some TX block if it exists, otherwise $\bot$.

\emph{Agreed enclosure} is the same as enclosure with an extra constraint where the CP blocks must be in some consensus result $\C_r$ and also must be the smallest enclosure.
That is, suppose a chain is in the form
    $\{c_{i}, c_{i+1}, t_{i+2}, c_{i+3}\}$\footnote{Usually the notation is of the form $c_{u, i}$, but the node identity is not important here so we simplify it to $c_{i}$}
    and $c_{i}, c_{i+1}, c_{i+3}$ are in $\C_r, \C_{r+1}, \C_{r+3}$ respectively,
    then the agreed enclosure of $t_{i+2}$ is $\langle c_{i+1}, c_{i+3}\rangle$ and cannot be $\langle c_{i}, c_{i+3}\rangle$.
Similarly, \emph{agreed fragment} has the same definition as fragment but using agreed enclosure.
We define its function to be $\textsf{agreed\_fragment}(\cdot)$ which we use later in the validation protocol (\Cref{sec:vd-protocol}).

% The length of the fragment is constrained by $L$,
% namely $\forall t |\textsf{fragment}(t)| \le L$.
% The purpose to prevent spam and encourage nodes to create more CP blocks.
% $L$ should be sufficiently high so that busy nodes are not hindered by it.

@@

Our consensus protocol runs on top of Extended TrustChain.
It is directly related to the creation of CP blocks.
The objectives of the protocol are to
    allow honest nodes always make progress (in the form of creating new CP blocks),
    compute correct consensus result in every round
    and have unbiased election of facilitators.
Concretely, we define the necessary properties as follows.
\begin{definition}
\label{def:consensus}
\textbf{\emph{Properties of the consensus protocol}}

$\forall r \in \mathbb{N}$, the following properties must be satisfied.
\begin{enumerate}
    \item \emph{Agreement}:
        If one honest node outputs a list of facilitators $\F_r$,
        every other node honest decides on $\F_r$
    \item \emph{Validity}:
        If any honest node outputs $\F_r$, then $|\F_r| > n$ and $|\F_r|$ must contain at least $n - t$ honest nodes.
    \item \emph{Fairness}:
        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
    \item \emph{Termination}:
        At the end of the round, every node outputs some $\F_r$.
\end{enumerate}
\end{definition}
These properties may look like Byzantine consensus properties (which we describe next in \Cref{sec:acs-background}) but they have some subtle differences.
Firstly, they are properties for every node in the network and not just the facilitators.
Secondly, they must be satisfied for all rounds because the whole system falls apart if one of the property cannot be satisfied in one of the rounds.

Before describing the protocol in detail,
we take a brief detour to give background on the asynchronous subset consensus.
This is the primary building block of our protocol.
% Starting with the bootstrap phase and then moving on to the actual consensus phase.

\subsection{Background on Asynchronous Subset Consensus}
\label{sec:acs-background}

The best way to explain asynchronous subset consensus (ACS) is to contrast it with the typical Byzantine consensus.
We adapt the description from~\cite[Chapter 17]{podc}.
%The Byzantine consensus problem is a classical problem is distributed systems literature.
%It is first described by Leslie Lamport as the Byzantine general's problem in 1982~\cite{lamport1982byzantine}.
%Byzantine consensus is described as follows (adapted from ).
\begin{definition}
\textbf{\emph{Byzantine Consensus}}

There are $n$ nodes, of which at most $t$ might experience Byzantine fault.
Node $i$ starts with an input value $v_i$.
The nodes must decide for one of  those values, satisfying the the following.
\begin{enumerate}
    \item \emph{Agreement:}
        If a correct node outputs $v$, then every node outputs $v$.
    \item \emph{Validity:}
        The decision value must be the input value of a node.
    \item \emph{Termination:}
        All correct nodes terminate in finite time.
\end{enumerate}
\end{definition}
A node under Byzantine fault means that it can have arbitrary behaviour.
For example not sending message or colluding with other Byzantine nodes to undermine the entire system.
Note that the decision is on a single value.
This is in contrast to ACS which we describe next.

@@

There are $n$ nodes, of which at most $t$ might experience Byzantine fault.
Node $i$ starts with an non-empty set of input values $C_i$.
The nodes must decide an output $C$, satisfying the following.
\begin{enumerate}
    \item \emph{Agreement:}
        If a correct node outputs $C$, then every node outputs $C$.
    \item \emph{Validity:}
        If any correct node outputs a set $C$,
        then $|C| \ge n - t$ and $C$ contains the input of at least $n - 2t$ nodes.
    \item \emph{Totality:}
        If $n - t$ nodes receive an input, then all correct nodes produce an output.
\end{enumerate}
\end{definition}

ACS has the nice property of censorship resiliance when compared to other consensus algorithms.
For instance, Hyperledger and Tendermint uses Practical Byzantine Fault Tolerance (PBFT)~\cite{castro1999practical} as their consensus algorithm.
In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectivly filter transactions.
In contract, every party in ACS are involved in the proposal phase,
and it guarantees that if $n - 2t$ parties propose the same transaction, then it must be in the agreed output.
Thus, if some value is submitted to at least $n - 2t$ nodes, it is guaranteed to be in the consensus result.
For a detailed description of ACS we refer to the HoneyBadgerBFT work~\cite{miller2016honey}.

The main drawback with ACS and also Byzantine consensus algorithms is the high message complexity.
Typically, such protocols have a message complexity of $O(n^2)$, where $n$ is the number of parties.
Hence, it may work with a small number of nodes,
but it is infeasible for blockchain systems where thousands of nodes are involved.

@@

\subsection{Bootstrap Phase}
\label{sec:bootstrap}
Now we have all the necessary information to describe our consensus protocol.
We begin with the bootstrap phase and then move onto the actual consensus phase.

Recall that facilitators are computed from the consensus result,
but the consensus result is agreed by the facilitators.
Thus we have a dependency cycle.
The goal of the bootstrap phase is to give us a starting point in the cycle.

To bootstrap, imagine that there is some bootstrap oracle, that initiates the code on every node.
The code satisfied all the properties in \Cref{def:consensus}.
Namely every node has the same set of valid facilitators $\F_1$ that are randomly choosen.
This concludes the bootstrap phase.

@@

\subsection{Relationship with $\alpha$-Synchroniser}

@@

    \State $txid', pk'_u, m' \gets t_{v, j}$
    \If{$m \ne m' \vee pk_u \ne pk'_u$}
        \State \Return \emph{invalid}
    \EndIf \Comment no tampering
    \State
